From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-07 Pi+/paper/equivalence.tex
Revise equivalence section

\begin{definition}
  \begin{align*}
    \evalt_{0} & : U \to \UPlus                                             \\
    \evalt_{1} & : (c : X \iso Y) \to \evalt_{0}(X) \iso \evalt_{0}(Y)      \\
    \evalt_{2} & : (\alpha : p \Iso q) \to \evalt_{1}(p) \Iso \evalt_{1}(q) \\
  \end{align*}
\end{definition}

@@

\begin{definition}
  \begin{align*}
    \evalh_{0} & : \UPlus \to \UHat                                                          \\
    \evalh_{1} & : (c : t_{1} \iso t_{2}) \to \evalh_{0}(t_{1}) \isoh \evalh_{0}(t_{2})      \\
    \evalh_{2} & : (\alpha : c_{1} \Iso c_{2}) \to \evalh_{1}(c_{1}) \Isoh \evalh_{1}(c_{2}) \\
  \end{align*}
\end{definition}

@@

\begin{definition}
  \begin{align*}
    \quoteh_{0} & : \UHat \to \UPlus                                                            \\
    \quoteh_{1} & : (p : X_{1} \isoh X_{2}) \to \quoteh_{0}(X_{1}) \iso \quoteh_{0}(X_{2})      \\
    \quoteh_{2} & : (\alpha : p_{1} \Isoh p_{2}) \to \quoteh_{1}(p_{1}) \Iso \quoteh_{1}(p_{2}) \\
  \end{align*}
\end{definition}

\begin{proposition}
  $\evalh/\quoteh$ give a symmetric monoidal biequivalence between $\PiPlusCat$ and $\PiHatCat$.
\end{proposition}

@@

\begin{definition}
  \begin{align*}
    \evalt_{0} & : \UHat \to \UFin                                                          \\
    \evalt_{1} & : (c : t_{1} \isoh t_{2}) \to \evalh_{0}(t_{1}) \id \evalh_{0}(t_{2})      \\
    \evalt_{2} & : (\alpha : c_{1} \Isoh c_{2}) \to \evalt_{1}(c_{1}) \id \evalt_{1}(c_{2}) \\
  \end{align*}
\end{definition}

@@

\begin{definition}
  \begin{align*}
    \quotet_{0} & : \UFin \to \UHat                                                            \\
    \quotet_{1} & : (p : X_{1} \id X_{2}) \to \quoteh_{0}(X_{1}) \isoh \quoteh_{0}(X_{2})      \\
    \quotet_{2} & : (\alpha : p_{1} \id p_{2}) \to \quoteh_{1}(p_{1}) \Isoh \quotet_{1}(p_{2}) \\
  \end{align*}
\end{definition}

@@

\begin{proposition}
  $\evalt/\quotet$ give a symmetric monoidal biequivalence between $\PiHatCat$ and $\UFin$.
\end{proposition}

@@

\begin{definition}
  \begin{align*}
    \normt_{0} & : U \to \UPlus                                        \\
    \normt_{0} & = \quoteh_{0} \comp \evalt_{0}                        \\
    \\
    \normt_{1} & : (c : X \iso Y) \to \normt_{0}(X) \iso \normt_{0}(Y) \\
    \normt_{1} & = \quoteh_{1} \comp \evalt_{1}                        \\
  \end{align*}
\end{definition}

>>>>

\begin{definition}
  \begin{align*}
    \evalt_{0} & : U \to \UPlus                                             \\
    \evalt_{1} & : (c : X \iso Y) \to \evalt_{0}(X) \iso \evalt_{0}(Y)      \\
    \evalt_{2} & : (\alpha : p \Iso q) \to \evalt_{1}(p) \Iso \evalt_{1}(q)
  \end{align*}
\end{definition}

@@

\begin{definition}
  \begin{align*}
    \evalp_{0} & : \UPlus \to \UHat                                                          \\
    \evalp_{1} & : (c : t_{1} \iso t_{2}) \to \evalp_{0}(t_{1}) \isoh \evalp_{0}(t_{2})      \\
    \evalp_{2} & : (\alpha : c_{1} \Iso c_{2}) \to \evalp_{1}(c_{1}) \Isoh \evalp_{1}(c_{2})
  \end{align*}
\end{definition}

@@

\begin{definition}
  \begin{align*}
    \quotep_{0} & : \UHat \to \UPlus                                                            \\
    \quotep_{1} & : (p : X_{1} \isoh X_{2}) \to \quotep_{0}(X_{1}) \iso \quotep_{0}(X_{2})      \\
    \quotep_{2} & : (\alpha : p_{1} \Isoh p_{2}) \to \quotep_{1}(p_{1}) \Iso \quotep_{1}(p_{2})
  \end{align*}
\end{definition}

\begin{proposition}
  $\evalp/\quotep$ give a symmetric monoidal biequivalence between $\PiPlusCat$ and $\PiHatCat$.
\end{proposition}

@@

\begin{definition}
  \begin{align*}
    \evalh_{0} & : \UHat \to \UFin                                                          \\
    \evalh_{1} & : (c : t_{1} \isoh t_{2}) \to \evalh_{0}(t_{1}) \id \evalh_{0}(t_{2})      \\
    \evalh_{2} & : (\alpha : c_{1} \Isoh c_{2}) \to \evalh_{1}(c_{1}) \id \evalh_{1}(c_{2})
  \end{align*}
\end{definition}

@@

\begin{definition}
  \begin{align*}
    \quoteh_{0} & : \UFin \to \UHat                                                            \\
    \quoteh_{1} & : (p : X_{1} \id X_{2}) \to \quoteh_{0}(X_{1}) \isoh \quoteh_{0}(X_{2})      \\
    \quoteh_{2} & : (\alpha : p_{1} \id p_{2}) \to \quoteh_{1}(p_{1}) \Isoh \quotet_{1}(p_{2})
  \end{align*}
\end{definition}

@@

\begin{proposition}
  $\evalh/\quoteh$ give a symmetric monoidal biequivalence between $\PiHatCat$ and $\UFin$.
\end{proposition}

The semantics that we presented here takes a different route to constructing the permutation from a $\PiLang$
combinator, compared to the direct interpretation given using the big-step interpreter in~\cref{subsec:denotational}. We
verify that the two semantics agree, establishing that the semantics is adequate and fully abstract.

\begin{definition}[${\gdenot{\blank}} : U \to {\UFin}$]
  \begin{align*}
    \gdenot{\blank}_{0} & = \evalh_{0} \comp \evalp_{0} \comp \evalt_{0} \\
    \gdenot{\blank}_{1} & = \evalh_{1} \comp \evalp_{1} \comp \evalt_{1}
  \end{align*}
\end{definition}

\begin{proposition}
  If $c_1, c_2 : X \iso Y$, then $\denot{c_1} = \denot{c_2}$ iff $\gdenot{c_1}_{1} = \gdenot{c_2}_{1}$.
\end{proposition}

@@

\begin{definition}
  \begin{align*}
    \normt_{0} & : U \to \UPlus                                            \\
    \normt_{0} & = \quotep_{0} \comp \quoteh_{0} \comp \gdenot{\blank}_{0} \\
    \\
    \normt_{1} & : (c : X \iso Y) \to \normt_{0}(X) \iso \normt_{0}(Y)     \\
    \normt_{1} & = \quotep_{1} \comp \quoteh_{1} \comp \gdenot{\blank}_{1}
  \end{align*}
\end{definition}
